Double White Dwarf Mergers on Adaptive Meshes\\ I. Methodology and Code Verification

### 2015-08-14 papers/wdmerger_I/paper.tex

We employ a ``sponge'' similar to that described by \citet{maestro3} for the outer regions of the 
computational domain. After the hydrodynamics update, we apply a damping force to the momentum 
equation as follows:
\begin{equation}
  (\rho \mathbf{u})^{n+1} \to \frac{(\rho \mathbf{u}^{n+1})}{1 + (\Delta t / \Delta t_S) f_S},
\end{equation} 
where $\Delta t_S$ is a timescale for the sponge to operate on, and $f_S$ is the damping factor.
We choose it so that that the sponge is non-operational inside a radius $r_S$ from the origin, and 
fully applied at a radius $r_S^\prime \equiv r_S + \Delta r_S$. We then smooth the sponge 
out between $r_S$ and $r_S^\prime$:
\begin{equation}
  f_{S} = \begin{dcases} 0 & r < r_S \\ \frac{1}{2}\left(1 - \text{cos}\left[\pi\left(\dfrac{r - r_S}{\Delta r_S}\right)\right]\right) & r_S \leq r < r_S^\prime \\ 1 & r \geq r_S^\prime. \end{dcases}
\end{equation}
For the simulations in this paper we set $r_S$ to be 75\% of the distance from the origin to the domain boundaries,
and $\Delta r_S$ so that the sponge extends another 10\% of that distance. We set 
$\Delta t_S = 0.01$ s, which is of the same order as the as the CFL timestep 
for typical problem setups. We find that this sponge is necessary especially in the 
rotating reference frame, as standing instabilities can create very large velocities 
in the ambient fluid that drag down the global timestep by up to an order of magnitude.
However, this provides another reason to distrust the dynamics of the low-density fluid
outside of the stellar orbits.

>>>>

We employ a ``sponge'' similar to that described by \citet{maestro3}
for the outer regions of the computational domain. After the
hydrodynamics update, we apply a damping force to the momentum
equation as follows:
\begin{equation}
  (\rho \mathbf{u})^{n+1} \to \frac{(\rho \mathbf{u}^{n+1})}{1 + (\Delta t / \Delta t_S) f_S},
\end{equation} 
where $\Delta t_S$ is a timescale for the sponge to operate on, and
$f_S$ is the damping factor.  We choose it so that that the sponge is
non-operational inside a radius $r_S$ from the origin, and fully
applied at a radius $r_S^\prime \equiv r_S + \Delta r_S$. We then
smooth the sponge out between $r_S$ and $r_S^\prime$:
\begin{equation}
  f_{S} = \begin{dcases} 0 & r < r_S \\ \frac{1}{2}\left(1 - \text{cos}\left[\pi\left(\dfrac{r - r_S}{\Delta r_S}\right)\right]\right) & r_S \leq r < r_S^\prime \\ 1 & r \geq r_S^\prime. \end{dcases}
\end{equation}
For the simulations in this paper we set $r_S$ to be 75\% of the \MarginPar{maybe a plot of the sponge profile?}
distance from the origin to the domain boundaries, and $\Delta r_S$ so
that the sponge extends another 10\% of that distance. We set $\Delta
t_S = 0.01$ s, which is of the same order as the as the CFL timestep
for typical problem setups. We find that this sponge is necessary
especially in the rotating reference frame, as standing instabilities
can create very large velocities in the ambient fluid that drag down
the global timestep by up to an order of magnitude.  However, this
provides another reason to distrust the dynamics of the low-density
fluid outside of the stellar orbits. \MarginPar{I am not sure I understand 
this statement}
